### 2017-01-27 CO750-ApproximationAlgorithmsTopics/lectures/lecture_07.tex
Reading

\lhead{2017-01-19}
\rhead{William Justin Toth CO750-Approximation Algorithms Lecture 5} %CHANGE n to ASSIGNMENT NUMBER ijk TO COURSE CODE
\renewcommand{\headrulewidth}{1pt} %heading underlined
%\renewcommand{\baselinestretch}{1.2} % 1.2 line spacing for legibility (optional)

@@

@@

Before moving on to the next topic, we just want to mention an important theorem that results from the theory built in the previous lectures. It tells us that we can have deterministic approximation algorithms that give just as good bounds as the randomized ones studied previously.

@@

Let $G=(V,E)$ be a graph. Let $d: E \rightarrow \R_+$ be a metric function. Let $\lambda^*_{uv} \geq 0$ be coefficients for all $(u,v) \in V\times V$. There exists a polynomial time algorithm which constructs a weighted tree $T=(V,E)$ such that $d^T(u,v) \geq d(u,v)$ for all $u,v\in V$ and $$\sum_{u,v} \lambda^*_{uv}d^T(u,v) \leq O(\log n) \sum_{u.v} \lambda^*_{uv} d(u,v).$$

@@

Omitted. May be in a talk at the end of course via some students project. Involves derandomization via the method of conditional expectation.

@@

\section{Primal - Dual Methods}

@@

Primal-Dual methods are a common paradigm in algorithm design. We will begin by showing how they are used to given exact algorithms, and then explain relaxed complementary slackness which allows us to extend the paradigm to give approximation algorithms.

@@

Consider the {\it Arborescence Problem} (ARB). In this problem we are given a directed graph $D= (V,E)$, cost function $c: E \rightarrow \R_+$, and a root vertex $r \in V$. We are to find an arborescence $T$ rooted at $r$ which spans $V$ and minimizes $\sum_{e\in E(T)} c_e$.
As the name of this method might suggest we will used integrally the linear programming relaxation of this problem:
\min \sum_{e \in E} c_e x_e  &\ \\
\text{s.t.} \sum_{e \in \delta^+(S)} x_e &\geq 1 &\forall S \subseteq V\backslash\{r\} \\
x_e &\geq 0 &\forall e \in E,
and its dual program:
\max \sum_{S \subseteq V\backslash\{r\}} y_S &\ \\
\text{s.t.} \sum_{S : e \in \delta^+(S)} y_S &\leq c_e &\forall e \in E \\
y_S &\geq 0 &\forall S \subseteq V\backslash \{r\}.
Here $\delta^+(S) = \{(u,v) \in E : u \in S, v\not\in S\}$ denotes set of arcs outgoing from $S$.
Our primal-dual algorithm for (ARB) operates as follows (start by setting $x=0$, $y=0$, for implementation track only non-negative $y$ to avoid exponential runtime): 
\item While $x$ is not feasible for the primal:
	\item Increase $y_S$ uniformly for all ``strongly connected components" (with respect to $x$) $S$ where $x(\delta^+(S)) = 0$ until some constraints for an edge $e$ becomes tight.
	\item Set $x_e=1$ for tight edge $e$ in previous step.
\item Deletion Step: Consider each edge $e$ where $x_e = 1$ in reverse order from the order they were set to $1$ and set $x_e = 0$ as long as this does not cause the resulting $x$ to become infeasible.
\item Return $x$.

@@

The previous algorithm is an exact polynomial time algorithm for (ARB).

@@

When the algorithm terminates we have that if $x_e > 0$ then $x_e = 1$  and so the dual constraint is tight, that is:
$$\sum_{S : e\in\delta^+(S)} y_S = c_e.$$
This is immediate from our construction of $x$. Hence the primal complementary slackness conditions hold.
For the dual complementary slackness conditions, suppose there exists $S$ such that $y_S > 0$, yet $\sum_{e\in \delta^+(S)} x_e \geq 2$. Let $e, \bar{e} \in \delta^+(S)$ and suppose without loss of generality that $\bar{e}$ was added before $e$ by the algorithm. Since we did not delete $e$, there exists a node $v \in S$ that uses $e$ to reach $r$. But since $S$ is strongly connected, no nodes in $S$ need $\bar{e}$ to reach $r$ (they can go $v$ then eventually through $e$), so we could have deleted $\bar{e}$ in the Deletion Step. A contradiction.
To extend this method to one which yields approximation algorithms, we need to be able to relax our complementary slackness conditions. In (ARB) our relaxed complementary slackness conditions are:  $x$ (with corresponding dual solution $y$) is an $\alpha \beta$-approximate solution if and only if
\item If $x_e > 0$ then $\sum_{S: e \in \delta^+(S)} y_S \geq \frac{1}{\alpha} c_e$.
\item If $y_S > 0$ then $\sum_{e\in \delta^+(S)} x_e \leq \beta \cdot 1$.
To see it observe that
$$\sum_{e} c_e x_e \leq \sum_e x_e(\alpha \sum_{S: e \in \delta^+(S)} y_S) = \alpha \sum_{e}x_e(\sum_{S: e \in \delta^+(S)} y_S) \leq \alpha \beta \sum_S y_S.$$
\subsection{Primal-Dual Algorithm for General Network Design Problems}
In a {\it general network design problem} (GND) we are given a graph $G = (V,E)$, a cost function $c : E \rightarrow \R_+$, and a function $f: 2^S \rightarrow \R_+$. We are to solve the integer program:
min \sum_{e \in E} c_e x_e  &\ \\
\text{s.t.} \sum_{e \in \delta(S)} x_e &\geq f(S) &\forall S \subseteq V\\
x_e &\geq 0 &\forall e \in E \\
x \in \Z^E.
This problem's linear programming relaxation is:
min \sum_{e \in E} c_e x_e  &\ \\
\text{s.t.} \sum_{e \in \delta(S)} x_e &\geq f(S) &\forall S \subseteq V \\
x_e &\geq 0 &\forall e \in E 
with corresponding dual program:
\max \sum_{S} f(S)y_S &\ \\
\text{s.t.} \sum_{S : e \in \delta(S)} y_S &\leq c_e &\forall e \in E \\
y_S &\geq 0 &\forall S \subseteq V.
We will focus our attention on the special case where $f$ is a $01$-proper function. A function $f$ is called {\it proper} if it satisfies:
\item $f(V) = 0$
\item $f(S) = f(V\backslash S)$, $\forall S\subseteq V$ (Symmetry property)
\item If $A \cap B = \emptyset$ then $f(A \cup B) \leq \max \{f(A), f(B)\}$ (Maximality property).
Our algorithm for (GSN) where $f$ is $01$-proper is as follows:
\item Start with ($x=0,y=0$) and $F = \emptyset$.
\item While $F$ is not feasible:
	\item Increase $y_S$ uniformly for "minimal violated cuts" (with respect to $F$) until a constraint corresponding to some edge $e$ becomes tight in the dual program.
	\item For such $e$ set $x_e = 1$ and $F = F \cup \{e\}$.
\item Deletion Step: Consider all edges in $F$ in reverse order they were added in Step $2$. If $F\backslash\{e\}$ is feasible then set $F = F\backslash\{e\}$ and set $x_e = 0$.
\item Return $F$.
Let $f$ be a $01$ function satisfying the Maximality property. Then
\item $F \subseteq E$ is feasible if and only if every connected component $C$ of $(V,F)$ satisfies $f(C) = 0$.
\item The minimal violated sets (cuts) are connected components $C$ that have $f(C) = 1$.
Let $S$ be a minimal violated set with $f(S) = 1$. If $(V,S)$ contains some connected component $A$ and $S\backslash A := B$ is not empty then by Maximality either $f(A) = 1$ or $f(B) = 1$. But then either $A$ or $B$ contradicts the minimality of $S$. Hence $S$ is a connected component and the lemma follows.
Our algorithm above is a $2$-approximation algorithm for (GND) when $f$ is a $01$-proper function.
Polynomial run time follows from our lemma, because it allows us to find minimal violated cuts in polynomial time by checking the value of $f$ on connected components.
For the relaxed complementary slackness conditions observe that if $e \in F$ then $\sum_{S : e \in \delta(S)} y_S = c_e$ by construction.  It remains to show that for $y_S > 0$ we have $\sum_{e\in\delta(S)}  x_e = |\delta(S) \cap F| \leq 2$. We will show that this bound holds on average. That is that
$$ \sum_{e \in F} c_e = \sum_{e \in F} (\sum_{S : e \in \delta(S)} y_S) = (a) = \sum_S |\delta_F(S)| y_S \leq 2 \sum_S y_S = (b) $$
Let $\Delta_i$ be the amount by which all minimum violated sets get increased in iteration $i$. Call such sets ACTIVE in iteration $i$ (denoted $A(i)$). Then
$$(a) = \sum_i \sum_{S \in A(i)} |\delta_F(S)| \Delta_i \quad (b) = \sum_i \sum_{S\in A(i)} \Delta_i.$$
We will show that for every iteration $i$, $\sum_i |\delta_F(S)| \leq 2 \cdot |A(i)|.$
So fix some $i$. Consider $(V,F)$ (which is a forest as cycles can be removed in Deletion Step), and contract all sets $S$ in $A(i)$ into one node. Call the resulting graph $H$.
\paragraph{Claim 1}
$H$ is a forest.
\paragraph{Proof of Claim 1}
Suppose the contraction of one set $S$ creates a cycle. But then in the reverse deletion step, one the two edges in $\delta_F(S)$ could be removed without violating feasibility, a contradiction. $\blacksquare$
\paragraph{Claim 2}
All leaves of $H$ correspond to active nodes.
\paragraph{Proof of Claim 2}
Let $c$ be a node of a component $\bar{H}$ of $H$. Suppose $f(c) = 0$. Note that $f(\bar{H}) = 0$ by our previous lemma. By symmetry $f(V\backslash \bar{H}) = 0$. By maximality $f((V\backslash \bar{H})\cup \{c\}) = 0$. By symmetry $f(\bar{H}\backslash \{c\}) = 0$. This implies that the edge connecting $c$ to $\bar{H}$ could have been removed in the deletion step, a contradiction.$\blacksquare$
Now our claims imply the following. Let $W$ be the set of shrunk nodes in $H$, and $\bar{W}$ the remains nodes. Discard isolated nodes in $H$. Then
$$\sum_{v\in W} d_H(v) + \sum_{v\in\bar{W}}d_H(v) \leq 2|W \cup \bar{W}|$$
since $H$ is a forest. But then $\sum_{v\in W} |\delta_H(v)| \leq 2|W|$, which is exactly equivalent to that which we set out to show. 

>>>>

\lhead{2017-01-26}
\rhead{William Justin Toth CO750-Approximation Algorithms Lecture n} %CHANGE n to ASSIGNMENT NUMBER ijk TO COURSE CODE
\renewcommand{\headrulewidth}{1pt} %heading underlined
%\renewcommand{\baselinestretch}{1.2} % 1.2 line spacing for legibility (optional)

@@

\section*{Integrality Gap}

@@

Let $P$ be an optimization problem, and suppose that there exists an integer programming formulation for $P$. Denote by $OPT(I)$ the objective function value of an optimal solution to an instance $I$ of $P$. Denote by $OPT_f(I)$ the objective function value of the Linear Programming relaxation of the IP formulation of $I$.
The integrality gap of the formulation is
$$\inf_{I \in P} \frac{OPT(I)}{OPT_f(I)}$$
for maximization problems, and
$$\sup_{I \in P} \frac{OPT(I)}{OPT_f(I)}$$
for minimization problems.
\section{Iterative Rounding}
\paragraph{}  Recall the linear programming formulation $LP$ for Generalized Steiner Network (GSN):
\min &\sum_e c_e x_e \\
\text{s.t.} x(\delta(S)) &\geq f(S) &\forall S \subseteq V\\
x&\geq 0
where $f(S) = \max_{u \in S, v \not\in S} r(u,v).$
A function $f: 2^V \rightarrow \Z$ is called weakly-supermodular (wsm) if
\item $f(V) = 0$ and
\item For all $A,B \subseteq V$ at least one of the following holds
\item $f(A) + f(B) \leq f(A\backslash B) + f(B \backslash A)$ or
\item $f(A) + f(B) \leq f(A\cup B) + f(A\cap B).$
\begin{proposition}
If $f$ is a proper function then $f$ is wsm.
\end{proposition}
Recall that an extreme point solution (xps) to an LP is a feasible solution which cannot be written as a convex combination of other feasible solutions.
\begin{proposition}
An optimal xps of $LP$ for GSN can be computed in polynomial time.
\end{proposition}

@@

For any wsm function $f$, any xps to $LP$ satisfies $x_e \geq \frac{1}{2}$ for at least one $e \in E$.
The previous theorem yields a natural algorithm for GSN:
\item Set $H= \emptyset$, and $f' = f$.
\item While $f' \neq 0$
\item Find an optimal xps for $LP$ with $f'$
\item For all $e$ such that $x_e \geq \frac{1}{2}$, include $\ceil{x_e}$ copies of $e$ in $H$
\item Update $f'$: for all $S\subseteq V$, set $f(S) = \max \{f(S)-|\delta_H(S)|, 0\}$.
\item Output $H$.
Let $H$ be a (multi)subgraph of $G$. If $f: 2^V \rightarrow \Z_+$ is wsm then $f'(S) = \max \{ f(S) - |\delta_H(S)|, 0\}$ is wsm.
We can solve the $LP$ relaxation for GSN in polynomial time at any iteration of the algorithm for instances of GSN.
The solution $H$ output by the algorithm has cost at most $2$ times the cost of the optimal solution to the $LP$ relaxation.

@@

Proceed by induction on the number of iterations. If the algorithm only requires one iteration then the result is immediate. Otherwise let $x$ be an optimal xps in the first iteration. Let $\hat{x}$ be defined as
$$ \hat{x}_e = \begin{cases} 1, &\forall e, x_e \geq \frac{1}{2} \\
0, &\text{otherwise}\end{cases}$$
Let $H^1$ be the set of edges taken in the first iteration. Then $c(H^1) \leq 2c(\hat{x})$. Let $f'$ be the requirement function after the first iteration and let $H'$ be the set of edges picked by the algorithm in all subsequent iterations. 
The key observation: $(x-\hat{x})$ is a feasible solution for $f'$. So by induction $c(H') \leq 2c(x-\hat{x})$. In total
$$c(H^1 \cup H') \leq 2c(\hat{x}) + 2c(x-\hat{x}) \leq 2c(x).$$

@@

@@

An xps $x$ of $LP$ satisfies $m = |E|$ linearly independent inequalities with equality (tight). In our analysis to follow we will make the simplifying assumption that $x_e > 0$ for all $e\in E$. We may assume this by removing all $e$ such that $x_e = 0$. For a set $S$ we will denote by $\chi(S)$ the incidence vector of $S$.

@@

A collection $\cL$ of subsets of $V$ is a \textit{laminar} family if no two sets in this collection ``crosses". By crosses we mean for all $A, B \in \cL$, we have $A\cap B \neq \emptyset$ and either $A\backslash B \neq \emptyset$ or $B\backslash A \neq \emptyset$.

@@

For any xps $x$ to $LP$ there is a collection of $m$ tight sets such that their incidence vectors are linearly independent, and the collection is a laminar family.

@@

To follow in next lecture.

### 2017-01-27 CS798-ConvexityAndOptimization/reading/reading_02.tex
Reading

\lhead{2017-01-26}
\rhead{William Justin Toth CO750-Approximation Algorithms Lecture n} %CHANGE n to ASSIGNMENT NUMBER ijk TO COURSE CODE
\renewcommand{\headrulewidth}{1pt} %heading underlined
%\renewcommand{\baselinestretch}{1.2} % 1.2 line spacing for legibility (optional)

\begin{document}
\section*{Integrality Gap}
\paragraph{}
Let $P$ be an optimization problem, and suppose that there exists an integer programming formulation for $P$. Denote by $OPT(I)$ the objective function value of an optimal solution to an instance $I$ of $P$. Denote by $OPT_f(I)$ the objective function value of the Linear Programming relaxation of the IP formulation of $I$.
\paragraph{}
The integrality gap of the formulation is
$$\inf_{I \in P} \frac{OPT(I)}{OPT_f(I)}$$
for maximization problems, and
$$\sup_{I \in P} \frac{OPT(I)}{OPT_f(I)}$$
for minimization problems.
\section{Iterative Rounding}
\paragraph{}  Recall the linear programming formulation $LP$ for Generalized Steiner Network (GSN):
\begin{align*}
\min &\sum_e c_e x_e \\
\text{s.t.} x(\delta(S)) &\geq f(S) &\forall S \subseteq V\\
x&\geq 0
\end{align*}
where $f(S) = \max_{u \in S, v \not\in S} r(u,v).$
\paragraph{}
A function $f: 2^V \rightarrow \Z$ is called weakly-supermodular (wsm) if
\begin{enumerate}
\item $f(V) = 0$ and
\item For all $A,B \subseteq V$ at least one of the following holds
\begin{enumerate}
\item $f(A) + f(B) \leq f(A\backslash B) + f(B \backslash A)$ or
\item $f(A) + f(B) \leq f(A\cup B) + f(A\cap B).$
\end{enumerate}
\end{enumerate}
\begin{proposition}
If $f$ is a proper function then $f$ is wsm.
\end{proposition}
\paragraph{}
Recall that an extreme point solution (xps) to an LP is a feasible solution which cannot be written as a convex combination of other feasible solutions.
\begin{proposition}
An optimal xps of $LP$ for GSN can be computed in polynomial time.
\end{proposition}
\begin{theorem}
For any wsm function $f$, any xps to $LP$ satisfies $x_e \geq \frac{1}{2}$ for at least one $e \in E$.
\end{theorem}
\paragraph{}
The previous theorem yields a natural algorithm for GSN:
\begin{enumerate}
\item Set $H= \emptyset$, and $f' = f$.
\item While $f' \neq 0$
\begin{enumerate}
\item Find an optimal xps for $LP$ with $f'$
\item For all $e$ such that $x_e \geq \frac{1}{2}$, include $\ceil{x_e}$ copies of $e$ in $H$
\item Update $f'$: for all $S\subseteq V$, set $f(S) = \max \{f(S)-|\delta_H(S)|, 0\}$.
\end{enumerate}
\item Output $H$.
\end{enumerate}
\begin{lemma}
Let $H$ be a (multi)subgraph of $G$. If $f: 2^V \rightarrow \Z_+$ is wsm then $f'(S) = \max \{ f(S) - |\delta_H(S)|, 0\}$ is wsm.
\end{lemma}
\begin{lemma}
We can solve the $LP$ relaxation for GSN in polynomial time at any iteration of the algorithm for instances of GSN.
\end{lemma}
\begin{theorem}
The solution $H$ output by the algorithm has cost at most $2$ times the cost of the optimal solution to the $LP$ relaxation.
\end{theorem}
\begin{proof}
Proceed by induction on the number of iterations. If the algorithm only requires one iteration then the result is immediate. Otherwise let $x$ be an optimal xps in the first iteration. Let $\hat{x}$ be defined as
$$ \hat{x}_e = \begin{cases} 1, &\forall e, x_e \geq \frac{1}{2} \\
0, &\text{otherwise}\end{cases}$$
Let $H^1$ be the set of edges taken in the first iteration. Then $c(H^1) \leq 2c(\hat{x})$. Let $f'$ be the requirement function after the first iteration and let $H'$ be the set of edges picked by the algorithm in all subsequent iterations. 
\paragraph{}
The key observation: $(x-\hat{x})$ is a feasible solution for $f'$. So by induction $c(H') \leq 2c(x-\hat{x})$. In total
$$c(H^1 \cup H') \leq 2c(\hat{x}) + 2c(x-\hat{x}) \leq 2c(x).$$
\end{proof}
\paragraph{}
An xps $x$ of $LP$ satisfies $m = |E|$ linearly independent inequalities with equality (tight). In our analysis to follow we will make the simplifying assumption that $x_e > 0$ for all $e\in E$. We may assume this by removing all $e$ such that $x_e = 0$. For a set $S$ we will denote by $\chi(S)$ the incidence vector of $S$.
\paragraph{}
A collection $\cL$ of subsets of $V$ is a \textit{laminar} family if no two sets in this collection ``crosses". By crosses we mean for all $A, B \in \cL$, we have $A\cap B \neq \emptyset$ and either $A\backslash B \neq \emptyset$ or $B\backslash A \neq \emptyset$.
\begin{theorem}
For any xps $x$ to $LP$ there is a collection of $m$ tight sets such that their incidence vectors are linearly independent, and the collection is a laminar family.
\end{theorem}
\begin{proof}
To follow in next lecture.
\end{proof}
\end{document}

>>>>

\lhead{2017-01-27}
\rhead{William Justin Toth CS798-Convexity and Optimization Reading 2} %CHANGE n to ASSIGNMENT NUMBER ijk TO COURSE CODE
\renewcommand{\headrulewidth}{1pt} %heading underlined
%\renewcommand{\baselinestretch}{1.2} % 1.2 line spacing for legibility (optional)

\begin{document}
\paragraph{}
We cover Sections 5.8 and 5.9 of Convex Optimization by Boyd and Vandenberghe.
\section*{5.8 - Theorems of  Alternatives}
\subsection*{Weak Alternatives via the dual}
\paragraph{}
We study the use of Lagrangian Duality Theory to the problem of determining feasibility for a set of inequality and equalities:
$$f_i(x) \leq 0,\ \ i=1,\dots,m \quad h_i(x) = 0, \ \ i=1,\dots,p $$
We assume the domain $\cD = \cap_{i=1}^m \textbf{dom} f_i \cap \cap_{i=1}^p \textbf{dom}h_i$ is nonempty. We can think of this as a standard minimization problem with objective function $0$ whose optimal value if $0$ if the system is feasible and $\infty$ otherwise.
\paragraph{}
We may associate with our inequality system a dual function
$$g(\lambda, \nu) = \inf_{x \in \cD} (\sum_{i=1}^m \lambda_i f_i(x) + \sum_{i=1}^p \nu_i h_i(x))$$
The associated dual problem to our associated primal problem is to maximize $g(\lambda, \nu)$ subject to $\lambda \geq 0$. Since $f_0 = 0$ this dual as optimal value
$$ d^* = \begin{cases}
\infty, &\lambda \geq 0 g(\lambda,\nu) > 0 \text{ is feasible} \\
0, &\text{otherwise}.
\end{cases}
$$
Hence at most one of our primal and dual systems is feasible since the dual lower bounds the primal. The two systems are called $\textit{weak alternatives}$ since they satisfy this property. Notice the primal system need not be convex, but the dual system always is.
\subsection*{Strong Alternatives}
\paragraph{}
When the original inequality system is convex, and some type of constraint qualification holds, we can obtain that the systems are $\textit{strong alternatives}$ which means that exactly one of them always holds.
\paragraph{}
Consider the system
$$f_i(x) \leq 0,\ \ i=1,\dots,m \quad Ax=b $$
and its alternative
$$\lambda \geq 0 \quad g(\lambda, \nu) > 0$$
Provided there exists an $x \in \textbf{relint} \cD$ with $Ax = b$ and optimal value $p^*$ (the minimum slack scalar optimization problem) that , then the two systems above are strong alternatives. With these assumptions we have strong duality $p^*=d^*$. Now suppose that our primal system is infeasible. Then $p^* > 0$. Then the dual optimal point $(\lambda^*, \nu^*)$ satisfies the alternative system.
\section*{5.9 - Generalized Inequalities}
\paragraph{}
The theory above associated with Lagrangian Duality can be extended to problems with generalized inequality constraints:
\begin{align*}
\min &f_0(x) \\
\text{s.t.} f_i(x) &\preccurlyeq_{K_i} 0, &i=1,\dots,m \\
h_i(x) & = 0, &i=1,\dots,p.
\end{align*}
where $K_i$ are proper cones (that is pointed closed convex cones). We need not assume the problem is convex, and as before we assume $\cD \neq \emptyset$. The order associated with cone $K_i$, denoted $\preccurlyeq_{K_i}$ is given as $A \preccurlyeq_{K_i} B$ if and only if $B-A \in K_i$.
\end{document}
